---
title: "Regression Inference - Tasks"
author: "Adam Soliman"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

1\. Copy the loading and cleaning code from slide 3 and run it.

```{r}
library(tidyverse)

star_df = read.csv("https://www.dropbox.com/s/bf1fog8yasw3wjj/star_data.csv?dl=1")
star_df = star_df[complete.cases(star_df),]
star_df = star_df %>%
    filter(star %in% c("small","regular") &
               grade == "k") %>% 
    mutate(small = (star == "small"))
```

2\. Generate the bootstrap distribution of $b_\textrm{small}$ based on 1,000 samples drawn from `star_df`.  
*Hint 1*: Use mutate to change the type of the `small` variable, and make it numeric.  
*Hint 2*: use the appropriate functions and arguments from the `infer` package so use the help pages.  
*Hint 3*: in `calculate` set `stat` to `slope`.

```{r}
library(infer)
bootstrap_distrib <- star_df %>% 
    mutate(small=as.numeric(small)) %>% 
    specify(response = math, explanatory = small) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "slope")
```

3\. Plot this simulated sampling distribution and compute mean and the standard error of $b_small$.

```{r}
bootstrap_distrib %>%
    ggplot(aes(x = stat)) +
    geom_histogram(boundary = 9, binwidth = 0.5, col = "white", fill = "#d90502") +
    labs(
        x = "Bootstrap sample slope estimate",
        y = "Frequency"
    ) +
    theme_bw(base_size = 14)
    
mean(bootstrap_distrib$stat)
sd(bootstrap_distrib$stat)
```


## Task 2

1\. Using the bootstrap distribution you generated in Task 1, compute the 95% confidence interval using the percentile method.

```{r}
ci_pctile = bootstrap_distrib %>%
    summarise(
        lower = quantile(stat, 0.025),
        upper = quantile(stat, 0.975)
    )
ci_pctile
```

2\. How similar is it to the confidence intervals obtained in the previous slide?

```{r}
# standard error method
ci_stderror <- bootstrap_distrib %>%
  summarise(
    lower = 8.895 - 1.96*sd(stat),
    upper = 8.895 + 1.96*sd(stat))
ci_stderror

# theory
library(broom)
ci_theory <- tidy(lm(math ~ small, star_df),
     conf.int = TRUE, conf.level = 0.95) %>%
  filter(term == "smallTRUE") %>%
  select(term, conf.low, conf.high)
ci_theory

bootstrap_distrib %>%
    ggplot(aes(x = stat)) +
    geom_histogram(boundary = 9, binwidth = 0.5, col = "white", fill = "#d90502") +
    labs(
        x = "Bootstrap sample slope estimate",
        y = "Frequency",
        title = "95% confidence interval computed with different methods",
        subtitle = "percentile (dashed), standard error (longdashed) and theory (solid)"
    ) +
  geom_vline(xintercept = c(ci_pctile$lower, ci_pctile$upper), linetype = "dashed", show.legend = TRUE) +
  geom_vline(xintercept = c(ci_stderror$lower, ci_stderror$upper), linetype = "longdash", show.legend = TRUE) +
  geom_vline(xintercept = c(ci_theory$conf.low, ci_theory$conf.high)) +
  theme_bw(base_size = 14)
```

