---
title: "Linear Regression Extensions - Tasks"
author: "Dr. Soliman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Task 1: Non-linear relationships

1\. Load the data [here](https://www.dropbox.com/scl/fi/68taafng8vjpzmub6q1xj/college_tuition_income.csv?rlkey=0c59a2kj0cwmcecjbg20jrkva&dl=0). This dataset contains information about tuition and estimated incomes of graduates for universities in the US.

```{r}
# change the path to yours
college <- read.csv("/Users/adamsoliman/Library/CloudStorage/Dropbox/Clemson/Econometrics Course/data for tasks/college_tuition_income.csv")
```


2\. Create a scatter plot of estimated mid career pay (`mid_career_pay`) $(y-axis)$ as a function of out of state tuition (`out_of_state_tuition`) $(x-axis)$. Would you say the relationship is broadly linear or rather non-linear? Use `geom_smooth(method = "lm", se = F) + geom_smooth(method = "lm", se = F, formula = y ~ poly(x, 2, raw= T))` to fit both a linear and 2nd order regression line. This time which seems most appropriate?

```{r}
base_graph <- ggplot(college, aes(x = out_of_state_tuition, y = mid_career_pay)) +
    geom_point() +
    labs(x = "Tuition for out-of-state residents in USD",
         y = "Estimated mid career pay in USD",
         title = "Relationship between out of state tuition and estimate mid career pay") +
    scale_x_continuous(labels = scales::dollar) +
    scale_y_continuous(labels = scales::dollar) +
    theme_bw(base_size = 14)
base_graph
```

**The relationship appears to be relatively linear though there is a some non-linearity at the upper end of the out of state tuition distribution, i.e. on the right-hand side of the graph. A 2nd order or 3rd order polynomial regression may be appropriate to capture it.**

```{r}
base_graph +
    geom_smooth(method='lm', se = FALSE, formula = y~poly(x,2, raw = T), aes(colour="y = b0 + b1x + b2x^2 + e")) +
    geom_smooth(method='lm', se = FALSE, aes(colour="y = b0 + b1x + e")) +
    scale_colour_manual(name = NULL, values=c("#d90502", "#DE9854")) +
    theme(legend.position = c(0,1),
          legend.justification = c(0,1))
```

**Based on my first impressions, the 2nd order polynomial fit seems to be a bit better at capturing non-linearities to the left and right of the graph.**

3\. Create a variable equal to out of state tuition divided by 1000. Regress mid career pay on out of state tuition divided by 1000. Interpret the coefficient.

```{r}
college <- college %>%
    mutate(out_state = out_of_state_tuition/1000)

lm(mid_career_pay ~ out_state, college)
```

**On average, a $1,000 increase in out of state tuition is associated with a $772 increase in estimated mid career pay.**

4\. Regress mid career pay on out of state tuition divided by 1000 and its square. *Hint:* you can use either `poly(x, 2, raw = T)` or `x + I(x^2)`, where x is your regressor. What does the positive sign on the squared term imply?

```{r}
lm(mid_career_pay ~ poly(out_state, 2, raw = T), college)
```

**The positive sign on the squared term (`poly(out_state, 2, raw = T)2` implies that as out of state tuition increases the increase in estimated mid career pay increases faster. In other words, the slope of the line is increasing in out of state tuition, it becomes steeper as you move to the right of the graph.**


## Task 2: Standardized regression

Let's go back our [grades](https://www.dropbox.com/scl/fi/77u2r8xznh63jg7aisq4f/grade5_isl.csv?rlkey=uclo7ng24so9fct41kcjm8k0l&dl=1) dataset. Please download it from the previous link and read it in using the appropriate function. Below are the estimates we got from regressing average math test scores on the full set of regressors.

```{r}
grades <- read.csv("/Users/adamsoliman/Library/CloudStorage/Dropbox/Clemson/Econometrics Course/data for tasks/grade5_isl.csv")

reg_full <- lm(avgverb ~ classize + disadvantaged + school_enrollment + female + religious, grades)
reg_full
```

1\. Create a new variable `avgmath_stand` equal to the standardized math score. You can use the `scale()` function (combined with `mutate`) or do it by hand with base `R`.

```{r}
# using scale()
grades <- grades %>%
    mutate(avgmath_stand = scale(avgmath))

# using base R
grades$avgmath_stand_2 <- (grades$avgmath - mean(grades$avgmath)) / sd(grades$avgmath)

identical(grades$avgmath_stand, grades$avgmath_stand_2)
```

2\. Run the full regression using the standardized math test score as the dependent variable. Interpret the coefficients and their magnitude.

```{r}
reg_full_stand <- lm(avgmath_stand ~ classize + disadvantaged + school_enrollment + female + religious, grades)
reg_full_stand$coefficients
```


**Interpretation for the class size coefficient: holding all the other regressors included in the regression constant, on average, a 1-student increase in class size is associated with a 0.003 standard deviation increase in average math scores. This is a very small association: 0.3% of a standard deviation.**

3\. Create the standardized variables for each *continuous* regressor as `<regressor>_stand`. 
  * Would it make sense to standardize the `religious` variable?

```{r}
grades <- grades %>%
    mutate(classize_stand = scale(classize),
           disadvantaged_stand = scale(disadvantaged),
           school_enrollment_stand = scale(school_enrollment),
           female_stand = scale(female))
```

**It would not make much sense to standardize the `religious` variable because it is a dummy variable, equal to 1 if the school is religious, and 0 otherwise.**

4\. Regress `avgmath_stand` on the full set of standardized regressors and `religious`. Discuss the relative influence of the regressors.

```{r}
reg_full_stand_2 <- lm(avgmath_stand ~ classize_stand + disadvantaged_stand + school_enrollment_stand + female_stand + religious, grades)
reg_full_stand_2$coefficients
```

**Interpretation for the class size coefficient: holding all the other regressors included in the regression constant, on average, a 1-standard deviation increase in class size (approx 6.5 students) is associated with a 0.02 standard deviation increase in average math scores. This is quite a small association: 2% of a standard deviation. Note that this coefficient can be otained by multiplying the class size coefficient from the previous question's regression by the approx 6.5 (i.e. 1 standard deviation of the class size variable).**

