---
title: "Introduction - Tasks"
author: "Dr. Soliman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

1\. Create a new R script (File $\rightarrow$ New File $\rightarrow$ R Script). Save it somewhere as `lecture_intro.R`.

2\. Type the following code in your script and run it. To run the code press `Ctrl` or `Cmd` + `Enter` (you can either highlight the code or just put your cursor at the end of the line)
```{r}
4 * 8
```

3\. Type the following code in your script and run it. What happens if you only run the first line of the code?
```{r}
x = 5 # equivalently x <- 5
x
```

```{r}
x = 5
```

**If I only run the first line of code, the object `x` is created in my environment but no output appears in the console. This is because I am not asking `R` to output anything; the only thing I am asking it is to create an object `x` equal to $5$.**

Congratulations, you have created your first `R` "object"! Everything is an object in R! Objects are assigned using `=` or `<-`.

4\. Create a new object named `x_3` to which you assign the cube of `x`. Note that to assign you need to use `=` or `<-`. Use code to compute the cube, not a calculator.

```{r}
x_3 = x^3
x_3
```

## Task 2

1\. Find out (using `help()` or google) how to import a .csv file. Do NOT use the "Import Dataset" button, nor install a package.

**There are many ways to import a .csv file. The simplest way is to use the `read.csv()` function.**

2\. Import [gun_murders.csv](https://www.dropbox.com/scl/fi/vanp2vdx741092v2k16dn/gun_murders.csv?rlkey=z81qu9y0f334s09538evukhh6&dl=1) in a new object `murders`. This file contains data on gun murders by US state in 2010. (Hint: objects are created using `=` or `<-`).

```{r}

# yours may look something like this:
murders = read.csv("/Users/adamsoliman/Library/CloudStorage/Dropbox/Clemson/Econometrics Course/data for tasks/gun_murders.csv")
# alternatively, you can download and read in directly from dropbox
murders <- read.csv("https://www.dropbox.com/scl/fi/vanp2vdx741092v2k16dn/gun_murders.csv?rlkey=z81qu9y0f334s09538evukhh6&dl=1")
```

3\. Ensure that `murders` is a data.frame by running:

```{r, error = TRUE}
class(murder)
```

**Hmm I get an error: "object 'murder' not found." Why is that? It's simple, there's a typo: it should be `murders` not `murder`. Everytime you see an error that says objet `x` not found it simply means that there is no object named `x` in your environment. So either you haven't created it yet or there's a small typo in your code. Always read what the error message says!**

```{r}
class(murders)
```

4\. Find out what variables are contained in `murders` by running

```{r}
names(murders)
```

5\. View the contents of `murders` by clicking on `murders` in your workspace. What does the `total` variable correspond to?

**The `total` variable corresponds to the total number of gun murders by state (in 2010).**

6\. How many observations are there in `murders`?

```{r}
nrow(murders)
```

7\. How many variables? What are the data types of each variable?

```{r}
str(murders)
```
