---
title: "Sampling - Tasks"
author: "Dr. Soliman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
```

## Task 

Instead of taking only 33 samples, let's take ***1000***!

1. Why do we not take 1000 samples "by hand"?

**It would be too time-consuming.**

2. Load the [data](https://github.com/adamsoliman/IntroEconometrics/blob/master/data%20for%20tasks/pasta.csv) into an object `pasta`.

```{r}
pasta <- read.csv("~/Documents/GitHub/Econometrics-Slides/data for tasks/pasta.csv")
```

3. Obtain 1000 samples of size 50 using the `rep_sample_n()` function from the `moderndive` package.

```{r}
virtual_samples <- pasta %>% 
    rep_sample_n(size = 50, reps = 1000)
```

4. Calculate the proportion of green pasta in each sample.

```{r}
virtual_prop_green <- virtual_samples %>% 
    group_by(replicate) %>% 
    summarize(
        num_green = sum(color == "green"),
        sample_n = n()) %>% 
    mutate(prop_green = num_green / sample_n)
```

5. Plot a histogram of the obtained proportion of green pasta in each sample.

```{r}
virtual_prop_green %>% ggplot(
    aes(x = prop_green)) +
    geom_histogram(
        binwidth = 0.02,
        boundary = 0.41,
        color = "white",
        fill = "darkgreen") +
    labs(x = "Proportion of green pasta in sample",
         y = "Frequency",
         title = "Distribution of 1000 samples of size 50") +
    theme_bw(base_size = 14)
```

6. What do you observe? Which proportions occur most frequently? How does the shape of the histogram compare to when we took only 33 samples?

**The distribution looks very close to a normal distribution. The proportions that occur most frequently are around 0.5. Compared to taking only 33 samples, the distribution looks significantly closer to a normal distribution.**

7. How likely is it that we sample 50 pasta of which less than 20% are green?

**It is extremely unlikely.**